### 2016-12-05 paper/main.tex

%%
%% Copyright 2007, 2008, 2009 Elsevier Ltd
%%
%% This file is part of the 'Elsarticle Bundle'.
%% ---------------------------------------------
%%
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%%
%% The list of all files belonging to the 'Elsarticle Bundle' is
%% given in the file `manifest.txt'.
%%

%% Template article for Elsevier's document class `elsarticle'
%% with numbered style bibliographic references
%% SP 2008/03/01
%%
%%
%%
%% $Id: elsarticle-template-num.tex 4 2009-10-24 08:22:58Z rishi $
%%
%%
\documentclass[preprint,12pt,3p]{elsarticle}

%% Use the option review to obtain double line spacing
%% \documentclass[preprint,review,12pt]{elsarticle}

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%% \documentclass[final,1p,times]{elsarticle}
%% \documentclass[final,1p,times,twocolumn]{elsarticle}
%% \documentclass[final,3p,times]{elsarticle}
%% \documentclass[final,3p,times,twocolumn]{elsarticle}
%% \documentclass[final,5p,times]{elsarticle}
%% \documentclass[final,5p,times,twocolumn]{elsarticle}

%% if you use PostScript figures in your article
%% use the graphics package for simple commands
%% \usepackage{graphics}
%% or use the graphicx package for more complicated commands
%% \usepackage{graphicx}
%% or use the epsfig package if you prefer to use the old commands
%% \usepackage{epsfig}

\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{mathrsfs}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{grffile}

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers after \end{frontmatter}.
%% \usepackage{lineno}

%% natbib.sty is loaded by default. However, natbib options can be
%% provided with \biboptions{...} command. Following options are
%% valid:

%%   round  -  round parentheses are used (default)
%%   square -  square brackets are used   [option]
%%   curly  -  curly braces are used      {option}
%%   angle  -  angle brackets are used    <option>
%%   semicolon  -  multiple citations separated by semi-colon
%%   colon  - same as semicolon, an earlier confusion
%%   comma  -  separated by comma
%%   numbers-  selects numerical citations
%%   super  -  numerical citations as superscripts
%%   sort   -  sorts multiple citations according to order in ref. list
%%   sort&compress   -  like sort, but also compresses numerical citations
%%   compress - compresses without sorting
%%
%% \biboptions{comma,round}

% \biboptions{}
\newcommand{\bx}{\mathbf{x}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\D}{\mathscr{D}}
\newcommand{\N}{\mathscr{N}}
\newcommand{\by}{\mathbf{y}}
\newcommand{\var}{\mathrm{var}}
\DeclareMathOperator*{\argmin}{\arg\!\min}

\newtheorem{conjecture}{Conjecture}
\newtheorem{theorem}{Theorem}
\newtheorem{experiment}{Experiment}

\journal{STAT 241A}

\begin{document}

\begin{frontmatter}

\title{A Novel Sampling Method Using Stochastic Gradient Descent}

\author{Yu Wang, 3031741324\\ Weixin Cai, 26954913}
%\address[label2]{Address Two\fnref{label4}}

\begin{abstract}
Stochastic gradient descent (SGD), under mild conditions, can be viewed as a Markov chain. This relationship has been pointed out in a variety of papers \cite{kushner2012stochastic, bach2014adaptivity}. Although SGD has been extensively studied, its property as a Markov chain seems to be largely overlooked. What is its stationary distribution? What is its mixing time? How does its mixing time relate to its convergence rate? Studying these properties could offer a novel perspective to understand SGD.

Currently, there aren't a lot of schemes available to analyzing SGD's mixing time. Traditional ways includes coupling, diameter bounds, and bottle-neck ratio etc. (see \emph{Reversible Markov chains and random walks on graphs} by Aldous et al.\cite{aldous2002reversible}, and \emph{Markov chains and mixing times} by Levin et al.\cite{levin2009markov}). However, these methods are designed for the Markov chain on a graph and might not be able to be applied to SGD directly.

Our proposal is to bridge the gap between classical Markov chain analysis and SGD. Specifically,  we would like to explore mixing time related issues and obtain some novel theoretical and empirical findings on the mixing time of SGD (If successful). A specific question to answer is whether lower mixing time will imply faster convergence rate. The hunch is that if the corresponding Markov chain has a low mixing time, SGD should converge fast. This seems to be a novel perspective to understand SGD. Duchi et al.\cite{duchi2012ergodic} shows how the finite-sample convergence rate (expected convergence \& high-probability convergence) of stochastic sub-gradient descent depends on the mixing time of the process. If we can bound the mixing time, subsequently, we will be able to understand the convergence of this class of optimization techniques. Theoretical analysis as well as empirical experiments will be carried out to further explore this issue.

Our group contains two people. The tentative plan is that Yu focuses on the theoretical exploration and Weixin focuses on the empirical experiments. The final report will be written together.
\end{abstract}

\begin{keyword}
%% keywords here, in the form: keyword \sep keyword
Stochastic gradient descent \sep Sampling \sep Metropolis-Hastings
%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)
\end{keyword}

\end{frontmatter}

%%
%% Start line numbering here if you want
%%
% \linenumbers
% ===================================================================================
\section{Introduction} % (fold)
\label{sec:introduction}

The posterior sampling method is used to great advantage by a board range of statistical methodologies, such as EM algorithm \citep{Dempster_Maximum_1977}. From the Bayesian point of view, E-step of the EM algorithm can be seen as calculating the posterior distribution of the parameters of interest \citep{Tanner_The_1987}. Another crucial arena of posterior sampling is Bayesian modeling. For example, \cite{Geman_Stochastic_1993} studied image restoration as a maximum a posteriori (MAP) problem on lattice graph. They proposed a Markov Random Field - Gibbs class of method for solving the posterior inference.

Typically, to implement such algorithms, one must be able to sample from the posterior distribution $p(\theta|y,z)$. In general, however, the posterior does not have closed form and sampling could be difficult and sometimes intractable. Various alternatives have been proposed such as Monte-Carlo method to approximate the posterior in EM algorithm \citep{Wei_A_1990}. \cite{Tanner_The_1987} considered data augmentation approach for calculating posterior. Analysis of posterior is studied in \cite{Rubin_The_1981} as a Bayesian Bootstrap approach.

To formalize the goal, the observed data $y$ follows a probability distribution indexed by $\theta$ : $p(y | \theta)$. We want to generate i.i.d. samples
\begin{equation}
\label{eq:sample}
x\mathop \sim\limits^{i.i.d.} p(\theta |y)
\end{equation}

Our approach to the problem (\ref{eq:sample}) is related to classical stochastic gradient descent (SGD) algorithms \citep{Robbins_A_1951,Polyak_Acceleration_1992}, where one assumes access to samples $y_1, ..., y_n$ from the distribution $p$ and performs gradient updates using $\nabla P(\theta ;y)$. \emph{This reduces to the randomized incremental subgradient method of \cite{Nedic_Incremental_2001} when $P$ is concentrated on a set of $n$ points, giving an objective of the form
$p(x) = \frac{1}{n}\sum\limits_{i = 1}^n {{p_i}(x)}$.} SGD is widely applied in statistical estimation and machine learning literature \citep{LeCun_Efficient_2012,Finkel_Efficient_2008}. Classical analysis of SGD focus on convergence in nature and generally do not provide result on its sampling distribution. It is not shown until recently that SGD, under mild conditions, can be viewed as a Markov chain \citep{kushner2012stochastic,bach2014adaptivity}. No previous work has analyzed SGD as sampling technique.

The main result of this paper is that performing stochastic gradient steps with specifically chosen step size, xxxxx results in a provably accurate and efficient sampling procedure for Bayesian posteriors. The convergence to the stationary distribution is governed by problem-dependent terms (namely the xxxx of the functions $P$, and xxxx) familiar from previous results on Markov chain [yyyyyy] as well as terms dependent on the rate at which xxxxxxx. Our xxx main theorems characterize the convergence of SGD in terms of xxxx parameter, which is xxxxx (in a sense we make precise later). In particular, we show that the stationary distribution is xxxxx with high probability.

The remainder of the paper is organized as follows. The next section contains a description of the algorithm we analyze and our main technical results. Following that in Section 3, we show that the classic stochastic gradient descent can be viewed as a special case of our algorithm. We expand on these results and provide an upper bound of the mixing time throughout Section 4, and give numerical examples to compare our result with other sampling methods like xxxxxx. We conclude with discussion in Section 5.

% section introduction (end)
% ===================================================================================
%% main text
\section{Main Theorems}

\begin{algorithm}
{\small
Given objective function $h(x)$, stepsize $\alpha$, noise level $\sigma^2$,
\begin{algorithmic}\caption{Stochastic gradient descent (oracle)}\label{Alg:SGD}
\STATE {\bf Initialize} $ \bx^0$ arbitrarily;
\WHILE{stopping criterion not satisfied}
\STATE draw $\epsilon^t$ from Gaussian distribution $\N(0, \sigma^2)$.
\STATE
$\bx^{k+1} \gets \bx^k - \alpha\nabla f(\bx^k) + \epsilon^t$;
\STATE $k\gets k+1$;
\ENDWHILE
\STATE return $\bx^k$.
\end{algorithmic}}
\end{algorithm}

Define $\D(\R^p)$ to be the family of the density functions of smooth random variables in $\R^p$, i.e.,

\begin{equation}
\D(\R^p)\triangleq \{p(\bx)\Big| p\in C^1(\R^p),~p(\bx) \geq 0,~ \int_\bx p(\bx) = 1.\}
\end{equation}
Consider any $p(\bx)\in \D(\R^p)$, we have the following theorem:

\begin{conjecture}
For any $p(\bx)\in \D(\R^p)$, it is the stationary distribution of iterates in Algorithm 1 with parameter $(-\log p(\bx), \epsilon, 2\epsilon)$.
\end{conjecture}
\begin{proof}
Let's verify this theorem by checking the detailed balance condition. Denote $g(\by|\bx)$ to be the density function of the transition, because $\by = \bx - \alpha \nabla f(\bx^k) + \epsilon^t$ and $\epsilon^t\sim \N(0, \sigma^2)$, we know
\begin{equation}
g(\by|\bx) = (2\pi \sigma^2)^{-p/2} \exp(- \frac{\|\by - \bx + \alpha \nabla h(\bx)\|^2}{2\sigma^2}).
\end{equation}
Then the detailed balance condition is
\begin{equation}\label{Eq:detailBalance}
p(\bx^{k+1}) g(\bx^{k}|\bx^{k+1}) = p(\bx^{k}) g(\bx^{k+1}|\bx^{k}).
\end{equation}
Since we will only consider $\bx^k$ and $\bx^{k+1}$ in the proof. For notation ease, we denote $\bx^{k+1}$ as $\bx^+$.
In order to prove \ref{Eq:detailBalance}, we would like to show the ratio between LHS and RHS to be 1. To be more specific, we have (Recall that $\sigma^2 = \alpha$)
\begin{eqnarray}
\frac{p(\bx^{+}) g(\bx^{k}|\bx^{+})}{p(\bx^{k}) g(\bx^{+}|\bx^{k})} = & \exp(-h(\bx^{+}) - \frac{\|\bx^{k} - \bx^+ + \alpha \nabla h(\bx^{+})\|^2}{2\alpha} + h(\bx^{k}) + \frac{\|\bx^{+} - \bx^k + \alpha \nabla h(\bx^k)\|^2}{2\alpha}).
\end{eqnarray}
Note that
\[h(\bx^+) - h(\bx^k) = \frac{\nabla h(\bx^k) + \nabla h(\bx^+)}{2} \cdot (\bx^+ - \bx^k) + O(\|\bx^+ - \bx^k\|^2),
\]
and
\begin{eqnarray}
 & \frac{\|\bx^{k} - \bx^+ + \alpha \nabla h(\bx^{+})\|^2}{2\alpha} - \frac{\|\bx^{+} - \bx^k + \alpha \nabla h(\bx^k)\|^2}{2\alpha} \\
 =& -\frac{\nabla h(\bx^k) + \nabla h(\bx^+)}{2} \cdot (\bx^+ - \bx^k) + \alpha/2 \|\nabla h(\bx^+)\|^2 - \alpha/2 \|\nabla h(\bx^+)\|^2.
\end{eqnarray}
Here, $\|\bx^+ - \bx^k\|$ will be roughly no bigger than $\alpha$ because $\bx^+ = \bx^k - \alpha \nabla h(\bx^k) + \epsilon^k$.  Thus it means $O(\|\bx^+ - \bx^k\|) = O(\alpha^2)$. What's more, $\alpha/2 \|\nabla h(\bx^+)\|^2 - \alpha/2 \|\nabla h(\bx^+)\|^2 = O(\alpha^2)$.
Thus we know
\begin{eqnarray}
\log \frac{p(\bx^{+}) g(\bx^{k}|\bx^{+})}{p(\bx^{k}) g(\bx^{+}|\bx^{k})} = &O(\alpha^2).
\end{eqnarray}
Therefore, we know the detailed balance equality holds approximately especially when $\alpha$ is small.

When $h$ is a quadratic function, the detailed balance holds exactly.
\end{proof}

Some empirical demonstration. How accurate is this approximation?

Since this is only a very rough deduction, we would like to examine the stationary distribution of SGD empirically for some special cases.

\begin{experiment}
Take the density function of $X$ to be $f(x) \propto \exp(- x^4)$, let $2\sigma^2 = \epsilon = 0.005$, $x_0 = 1$, burning-in period is $1000$, waiting period is $1000$, the empirical histogram v.s. the desired histogram is plotted in Figure ???
\end{experiment}
\begin{figure}
    \centering
    \includegraphics[width=.3\textwidth]{../figure/case1_step_0.005_iter_1e3.png}
    \includegraphics[width=.3\textwidth]{../figure/case1_step_0.005_iter_1e4.png}
    \includegraphics[width=.3\textwidth]{../figure/case1_step_0.005_iter_1e5.png}
    \includegraphics[width=.3\textwidth]{../figure/case2_step_0.005_iter_1e3.png}
    \includegraphics[width=.3\textwidth]{../figure/case2_step_0.005_iter_1e4.png}
    \includegraphics[width=.3\textwidth]{../figure/case2_step_0.005_iter_1e5.png}
    \includegraphics[width=.3\textwidth]{../figure/case3_step_0.005_iter_1e3.png}
    \includegraphics[width=.3\textwidth]{../figure/case3_step_0.005_iter_1e4.png}
    \includegraphics[width=.3\textwidth]{../figure/case3_step_0.005_iter_1e5.png}
    \caption{Experiment 1: the empirical histogram using SGD. Each column corresponds to a different number of samples ($1e3$, $1e4$, and $1e5$), each row corresponds to a different distrubtion ($f_1(x) = \exp(-x^4)$, $f_2(x) = \exp(-(x^2 - 4)(x^2 - 1))$, $f_3(x) = \exp(-|x|)$}
    \label{Fig:exp1}
\end{figure}
% ===================================================================================
\section{Sampling from posterior distribution}
Consider a statistical model fitting problem. Given $n$ samples $X_1,\ldots, X_n$ and an assumed model $P(X_1|\theta)$. The maximum likelihood estimate will be
\[
\hat{\theta} \in \argmin - \sum_{i=1}^n \log P(X_i|\theta).
\]
This can be solved by gradient descent on objective function $- \sum_{i=1}^n \log P(X_i|\theta)$. Now let's consider a Bayesian approach to estimate $\theta$ by maximizing its posterior probability given the data. The posterior distribution of $\theta$ is
\[
P(\theta|(X_1,\ldots, X_n)) = \frac{\prod_{i=1}^n P(X_i|\theta) P(\theta)}{P(X_1,\ldots, X_n)} \propto \prod_{i=1}^n P(X_i|\theta) P(\theta).
\]
By Conjecture 1, we could use SGD to draw samples using $(-\sum_i \log P(X_i|\theta) + \log P(\theta), \epsilon, 2\epsilon)$.  Note that the core update rule is
\[
\theta^{t+1} \gets \theta^{t} + \alpha (n^{-1} \sum_i \nabla \log P(X_i|\theta^t) - \log P(\theta^t)) + \epsilon^t.
\]
Note that $\alpha (n^{-1} \sum_{i=1}^n \nabla \log P(X_i|\theta^t) - \log P(\theta^t)) + \epsilon^t$ can be approximated by
\[
\sum_{i=1}^K (K^{-1} \nabla \log P(X_{\eta_i}|\theta^t) - \log P(\theta^t)),
\]
where $K$ is a constant smaller than $n$ and $\eta_i$ are IID drawn from $1,\ldots, n$. When $K\ll n$, this is much computational efficient. This relationship is elaborated in the following theorem:
\begin{theorem}[Central limit theorem]
Let $K = \alpha^{-1} \var \log P(X_1|\theta^t)$,
\[
\sum_{i=1}^K (K^{-1}\nabla \log P(X_{\eta_i}|\theta^t) - \log P(\theta^t)) - (n^{-1} \sum_{i=1}^n \nabla \log P(X_i|\theta^t) - \log P(\theta^t)) \sim \N(0, \alpha).
\]
\end{theorem}
\begin{algorithm}
{\small
Given data $X_1,\ldots, X_n$, model $p(X|\theta)$, sample size $K$
\begin{algorithmic}\caption{SGD based sampling}\label{Alg:SGD_Sampling}
\STATE {\bf Initialize} $ \theta^0$, $t \gets 0$;
\WHILE{stopping criterion not satisfied}
\STATE draw $K$ iid samples with replacement from $1,\ldots, n$, denoted $\eta_1,\ldots, \eta_K$.
\STATE calculate the mean $a^t$ and the covariance matrix $\Sigma^t$ of $(\nabla\log p(X_{\eta_i}|\theta^t))_{i=1}^K$.
\STATE
$\theta^{t+1} \gets \theta^t + 2(\Sigma^t)^{-1}a^t$;
\STATE $t\gets t+1$;
\ENDWHILE
\STATE return $\theta^k$.
\end{algorithmic}}
\end{algorithm}

A toy example showing SGD works.
% ===================================================================================
\section{Upper bound on the mixing time} % (fold)
\label{sec:upper_bound_on_the_mixing_time}
In order to analyze the efficiency of SGD, we need to consider its mixing time. If SGD's mixing time is small, we know it will be a very efficient sampling method. How to estimate its mixing time? In the past literature, the upperbound of a mixing time can be obtained by a coupling technique. Define $\|\cdot\|_{TV}$ to be the total variation norm of $\cdot$, $d(t) = \max_{x\in \Omega} \|P^t(x, \cdot), \pi(\cdot)\|_{TV}$, and $\bar{d}(t) = \max_{x,y\in \Omega} \|P^t(x, \cdot), P^t(y,\cdot)\|_{TV}$. Define $t_{mix} = \min\{t:d(t) < \epsilon\}$. 
\begin{conjecture}[Upperbound for mixing time] $\tau$ is the convergence stoppiing time. $\tau_{mix} \leq 2 \max_{x} \E_{x}(\tau)$.
\end{conjecture}
% section upper_bound_on_the_mixing_time (end)
% ===================================================================================
\section{Numerical experiments} % (fold)
\label{sec:numerical_experiments}

$f_1(x) = \exp(-x^4)$

$f_2(x) = \exp(-(x^2 - 4)(x^2 - 1))$

$f_3(x) = \exp(-|x|)$

% section numerical_experiments (end)
% ===================================================================================
\section{Discussion} % (fold)
\label{sec:discussion}

- what we have achieved

- Future work

% section discussion (end)
\newpage
% ===================================================================================
%% The Appendices part is started with the command \appendix;
%% appendix sections are then done as normal sections
\appendix

\section{Section in Appendix}
\label{appendix-sec1}

Sample text. Sample text. Sample text. Sample text. Sample text. Sample text.
Sample text. Sample text. Sample text. Sample text. Sample text. Sample text.
Sample text.

%% References
%%
%% Following citation commands can be used in the body text:
%% Usage of \cite is as follows:
%%   \cite{key}         ==>>  [#]
%%   \cite[chap. 2]{key} ==>> [#, chap. 2]
%%

%% References with bibTeX database:

\bibliographystyle{elsarticle-num}
% \bibliographystyle{elsarticle-harv}
% \bibliographystyle{elsarticle-num-names}
% \bibliographystyle{model1a-num-names}
% \bibliographystyle{model1b-num-names}
% \bibliographystyle{model1c-num-names}
% \bibliographystyle{model1-num-names}
% \bibliographystyle{model2-names}
% \bibliographystyle{model3a-num-names}
% \bibliographystyle{model3-num-names}
% \bibliographystyle{model4-names}
% \bibliographystyle{model5-names}
% \bibliographystyle{model6-num-names}

\bibliography{sample}

\end{document}

%%
%% End of file `elsarticle-template-num.tex'.