### 2016-12-05 poster/main.tex
poster initial

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Jacobs Landscape Poster
% LaTeX Template
% Version 1.0 (29/03/13)
%
% Created by:
% Computational Physics and Biophysics Group, Jacobs University
% https://teamwork.jacobs-university.de:8443/confluence/display/CoPandBiG/LaTeX+Poster
%
% Further modified by:
% Nathaniel Johnston (nathaniel@njohnston.ca)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[final]{beamer}

% \usepackage[scale=1.24]{beamerposter} % Use the beamerposter package for laying out the poster
\usepackage[scale=0.8]{beamerposter}

\usetheme{confposter} % Use the confposter theme supplied with this template

\setbeamercolor{block title}{fg=ngreen,bg=white} % Colors of the block titles
\setbeamercolor{block body}{fg=black,bg=white} % Colors of the body of blocks
\setbeamercolor{block alerted title}{fg=white,bg=dblue!70} % Colors of the highlighted block titles
\setbeamercolor{block alerted body}{fg=black,bg=dblue!10} % Colors of the body of highlighted blocks
% Many more colors are available for use in beamerthemeconfposter.sty

%-----------------------------------------------------------
% Define the column widths and overall poster size
% To set effective sepwid, onecolwid and twocolwid values, first choose how many columns you want and how much separation you want between columns
% In this template, the separation width chosen is 0.024 of the paper width and a 4-column layout
% onecolwid should therefore be (1-(# of columns+1)*sepwid)/# of columns e.g. (1-(4+1)*0.024)/4 = 0.22
% Set twocolwid to be (2*onecolwid)+sepwid = 0.464
% Set threecolwid to be (3*onecolwid)+2*sepwid = 0.708

\newlength{\sepwid}
\newlength{\onecolwid}
\newlength{\twocolwid}
\newlength{\threecolwid}
% \setlength{\paperwidth}{48in} % A0 width: 46.8in
% \setlength{\paperheight}{36in} % A0 height: 33.1in
\setlength{\paperwidth}{36in} % A0 width: 46.8in
\setlength{\paperheight}{24in} % A0 height: 33.1in
\setlength{\sepwid}{0.024\paperwidth} % Separation width (white space) between columns
% \setlength{\onecolwid}{0.22\paperwidth} % Width of one column
% \setlength{\twocolwid}{0.464\paperwidth} % Width of two columns
% \setlength{\threecolwid}{0.708\paperwidth} % Width of three columns
\setlength{\onecolwid}{0.3013\paperwidth} % Width of one column
\setlength{\twocolwid}{0.6026\paperwidth} % Width of two columns
\setlength{\threecolwid}{0.904\paperwidth} % Width of three columns
\setlength{\topmargin}{-0.5in} % Reduce the top margin size
%-----------------------------------------------------------

\usepackage{graphicx}  % Required for including images
\usepackage{booktabs} % Top and bottom rules for tables
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{mathrsfs}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{grffile}

\newcommand{\bx}{\mathbf{x}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\D}{\mathscr{D}}
\newcommand{\N}{\mathscr{N}}
\newcommand{\by}{\mathbf{y}}
\newcommand{\var}{\mathrm{var}}
\DeclareMathOperator*{\argmin}{\arg\!\min}

\newtheorem{conjecture}{Conjecture}
% \newtheorem{theorem}{Theorem}
\newtheorem{experiment}{Experiment}

% inline theorem
\makeatletter
\setbeamertemplate{theorem begin}
  {\usebeamercolor[fg]{thcolor}% for the heading
  {\bfseries\inserttheoremname~}%
  \ifx\inserttheoremaddition\@empty\else(\inserttheoremaddition)\ \fi%
  \hspace{.01em}\normalfont\usebeamercolor[fg]{thcolor}% for the body
  }
\setbeamertemplate{theorem end}{}
\makeatother
%----------------------------------------------------------------------------------------
%	TITLE SECTION
%----------------------------------------------------------------------------------------

\title{Novel Sampling Using Stochastic Gradient Descent} % Poster title

\author{Yu Wang and Weixin Cai} % Author(s)

\institute{Department of Statistics, UC Berkeley} % Institution(s)

%----------------------------------------------------------------------------------------

\begin{document}

\addtobeamertemplate{block end}{}{\vspace*{2ex}} % White space under blocks
\addtobeamertemplate{block alerted end}{}{\vspace*{2ex}} % White space under highlighted (alert) blocks

\setlength{\belowcaptionskip}{2ex} % White space under figures
\setlength\belowdisplayshortskip{2ex} % White space under equations

\begin{frame}[t] % The whole poster is enclosed in one beamer frame

\begin{columns}[t] % The whole poster consists of three major columns, the second of which is split into two columns twice - the [t] option aligns each column's content to the top

\begin{column}{\sepwid}\end{column} % Empty spacer column

\begin{column}{\onecolwid} % The first column

%----------------------------------------------------------------------------------------
%	OBJECTIVES
%----------------------------------------------------------------------------------------

\begin{alertblock}{Main results}

% \centerline{ \bf{ Aim of study}}

\begin{enumerate}
\item Stochastic gradient descent (SGD) as a class of efficient Markov chain sampler
\item Proved stationary distribution
\item Fast sampler for any Bayesian posteriors
\item Derived upper bound on the mixing time based on convergence results
\end{enumerate}

\end{alertblock}

%----------------------------------------------------------------------------------------
%   MOTIVATION
%----------------------------------------------------------------------------------------

\begin{block}{Motivation}

\begin{itemize}
    \item aaa
\end{itemize}

\end{block}

%------------------------------------------------
%
%----------------------------------------------------------------------------------------
%	INTRODUCTION
%----------------------------------------------------------------------------------------

\begin{block}{SGD Sampling Algorithm}

\begin{algorithm}[H]
{\small
Given objective function $h(x)$, stepsize $\alpha$, noise level $\sigma^2$,
\begin{algorithmic}\caption{Stochastic gradient descent (oracle)}\label{Alg:SGD}
\STATE {\bf Initialize} $ \bx^0$ arbitrarily;
\WHILE{stopping criterion not satisfied}
\STATE draw $\epsilon^t$ from Gaussian distribution $\N(0, \sigma^2)$.
\STATE
$\bx^{k+1} \gets \bx^k - \alpha\nabla f(\bx^k) + \epsilon^t$;
\STATE $k\gets k+1$;
\ENDWHILE
\STATE return $\bx^k$.
\end{algorithmic}}
\end{algorithm}

\end{block}

%----------------------------------------------------------------------------------------

\end{column} % End of the first column

\begin{column}{\sepwid}\end{column} % Empty spacer column

\begin{column}{\onecolwid} % Begin a column which is two columns wide (column 2)

%------------------------------------------------

\begin{figure}
    \centering
    \includegraphics[width=.3\onecolwid]{../figure/case1_step_0.005_iter_1e3.png}
    \includegraphics[width=.3\onecolwid]{../figure/case1_step_0.005_iter_1e4.png}
    \includegraphics[width=.3\onecolwid]{../figure/case1_step_0.005_iter_1e5.png}\\
    \includegraphics[width=.3\onecolwid]{../figure/case2_step_0.005_iter_1e3.png}
    \includegraphics[width=.3\onecolwid]{../figure/case2_step_0.005_iter_1e4.png}
    \includegraphics[width=.3\onecolwid]{../figure/case2_step_0.005_iter_1e5.png}\\
    \includegraphics[width=.3\onecolwid]{../figure/case3_step_0.005_iter_1e3.png}
    \includegraphics[width=.3\onecolwid]{../figure/case3_step_0.005_iter_1e4.png}
    \includegraphics[width=.3\onecolwid]{../figure/case3_step_0.005_iter_1e5.png}
    \caption{Experiment 1: the empirical histogram using SGD. Each column corresponds to a different number of samples ($1e3$, $1e4$, and $1e5$), each row corresponds to a different distrubtion ($f_1(x) = \exp(-x^4)$, $f_2(x) = \exp(-(x^2 - 4)(x^2 - 1))$, $f_3(x) = \exp(-|x|)$}
    \label{Fig:exp1}
\end{figure}

%----------------------------------------------------------------------------------------
%	MATERIALS
%----------------------------------------------------------------------------------------

\begin{block}{Stationary Distribution}

\begin{conjecture}
For any $p(\bx)\in \D(\R^p)$, it is the stationary distribution of iterates in Algorithm 1 with parameter $(-\log p(\bx), \epsilon, 2\epsilon)$.
\end{conjecture}

\end{block}

%----------------------------------------------------------------------------------------
%   ADDITIONAL INFORMATION
%----------------------------------------------------------------------------------------

\begin{block}{SGD As Posterior Sampler}

\begin{theorem}[Central limit theorem]
Let $K = \alpha^{-1} \var \log P(X_1|\theta^t)$,
\[
\sum_{i=1}^K (K^{-1}\nabla \log P(X_{\eta_i}|\theta^t) - \log P(\theta^t)) - (n^{-1} \sum_{i=1}^n \nabla \log P(X_i|\theta^t) - \log P(\theta^t)) \sim \N(0, \alpha).
\]
\end{theorem}

A toy example showing SGD works.

\end{block}

%----------------------------------------------------------------------------------------

\end{column} % End of the second column

\begin{column}{\sepwid}\end{column} % Empty spacer column

\begin{column}{\onecolwid} % The third column

%------------------------------------------------

\begin{figure}
\includegraphics[width=0.4\linewidth]{../figure/simulation2_empirical.png}
\includegraphics[width=0.4\linewidth]{../figure/simulation2_real.png}
\caption{Figure caption}
\end{figure}

%----------------------------------------------------------------------------------------
%   ADDITIONAL INFORMATION
%----------------------------------------------------------------------------------------

\begin{block}{Upper Bound On Mixing Time}

Maecenas ultricies feugiat velit non mattis. Fusce tempus arcu id ligula varius dictum.
\begin{itemize}
\item Curabitur pellentesque dignissim
\item Eu facilisis est tempus quis
\item Duis porta consequat lorem
\end{itemize}

\end{block}

%----------------------------------------------------------------------------------------
%	CONCLUSION
%----------------------------------------------------------------------------------------

\begin{block}{Conclusion}

Nunc tempus venenatis facilisis. \textbf{Curabitur suscipit} consequat eros non porttitor. Sed a massa dolor, id ornare enim. Fusce quis massa dictum tortor \textbf{tincidunt mattis}. Donec quam est, lobortis quis pretium at, laoreet scelerisque lacus. Nam quis odio enim, in molestie libero. Vivamus cursus mi at \textit{nulla elementum sollicitudin}.

\end{block}

%----------------------------------------------------------------------------------------
%   ACKNOWLEDGEMENTS
%----------------------------------------------------------------------------------------

\begin{block}{Acknowledgements}

\small{\rmfamily{Nam mollis tristique neque eu luctus. Suspendisse rutrum congue nisi sed convallis. Aenean id neque dolor. Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas.}} \\

\end{block}

%----------------------------------------------------------------------------------------
%	REFERENCES
%----------------------------------------------------------------------------------------

\begin{block}{References}

\nocite{*} % Insert publications even if they are not cited in the poster
\small{\bibliographystyle{unsrt}
\bibliography{sample}\vspace{0.75in}}

\end{block}

%----------------------------------------------------------------------------------------
%	CONTACT INFORMATION
%----------------------------------------------------------------------------------------

\begin{block}{Contant Information}

\begin{itemize}
\item Email: \href{mailto:wang.yu@berkeley.edu}{wang.yu@berkeley.edu}
\item Email: \href{mailto:wcai@berkeley.edu}{wcai@berkeley.edu}
\end{itemize}

\end{block}

%----------------------------------------------------------------------------------------

\end{column} % End of the third column

\end{columns} % End of all the columns in the poster

\end{frame} % End of the enclosing frame

\end{document}